---
title: "p8105_hw6_jys2137"
author: "jys2137"
date: "12/4/2021"
output: github_document
---
```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(modelr)
library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
	dpi = 300,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

### Problem 1

In this problem, we analyze data gathered to understand the effects of several variables on a child’s birthweight. The `birthweight` dataset consists of roughly 4000 children and includes the following variables:

#### 1.1. Loading and cleaning  the data  

First, we load and clean the data in preparation for regression analysis. We do this by:

- converting _**numeric**_ variables to _**factor**_  (for `babysex`, `frace`, `malform`, and `mrace`)
- converting **imperial** variable measurements to **metric** for consistency (weight from pounds to grams, height from inches to centimeters).
- checking for **missing data**

```{r load and clean data}
birthweight_df = read_csv("data/birthweight.csv") %>% 
  janitor::clean_names() %>%
  mutate(
    babysex = as.factor(babysex),
    babysex = fct_recode(babysex, "male" = "1", "female" = "2"),
    frace = as.factor(frace),
    frace = fct_recode(frace, "white" = "1", "black" = "2", "asian" = "3", 
                       "puerto rican" = "4", "other" = "8"),
    mrace = as.factor(mrace),
    mrace = fct_recode(mrace, "white" = "1", "black" = "2", "asian" = "3", 
                       "puerto rican" = "4"),
    malform = as.factor(malform),
    malform = fct_recode(malform, "absent" = "0", "present" = "1"),
    delwt_g = delwt * 453.59237,
    mheight_cm = mheight * 2.54,
    ppwt_g = ppwt * 453.59237,
    wtgain_g = wtgain * 453.59237)

sum(is.na(birthweight_df))
```

Note that in our check for missing data, we can see that there are **no `NA` observations** in the data frame.


### Problem 2

For this problem, we explore boostrapping by using weather data from NOAA that involves the `minimum` and `maximum` temperatures for Central Park in 2017.

#### 2.1. Loading the data

The code chunk below loads the weather dataset.

```{r weather_df, cache = TRUE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

#### 2.2. Bootstrap
Now, we focus on a simple linear regression and are interested in the distribution of two quantities estimated from these data:

1. $\hat{r}^2$
2. $\log(\beta_0 * \beta1)$

We will use 5000 bootstrap samples from the `weather_df` data frame. For each sample, we fit a simple linear regression with `tmax` as the response and `tmin` as the predictor.

The chunk below sets the bootstraps up. We use `broom::glance()` for getting $\hat{r}^2$ and `broom::tidy()` for getting $\log(\beta_0 * \beta_1)$ from the fitted regression.

```{r bootstrap setup, warning = FALSE}
set.seed(123)

boot_strap =
  weather_df %>% 
  modelr::bootstrap(n = 5000) %>%
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x)),
    results = map(models, broom::tidy),
    glance = map(models, broom::glance)) %>% 
  select(-strap, -models) %>% 
  unnest(results, glance) %>% 
  select(.id, term, estimate, r.squared) %>% 
  pivot_wider(
    names_from = term,
    values_from = estimate) %>% 
  rename(
    beta0 = `(Intercept)`,
    beta1 = tmin) %>% 
  mutate(log_b0b1 = log(beta0 * beta1)) %>% 
  select(r.squared, log_b0b1)
```

##### 2.3. Estimate $\hat{r}^2$

To plot the distribution of the estimates of $\hat{r}^2$, we use the following code chunk. 

```{r boostrap_r_sq}
boot_strap %>% 
  ggplot(aes(x = r.squared)) + 
  geom_density(alpha = 0.2) +
  stat_function(fun = dnorm,
                args = with(boot_strap, c(mean = mean(r.squared), sd = sd(r.squared))),
                color = "red",
                size = 1.5,
                alpha = .7) +
  theme(legend.position = "none") +
  labs(
        title = "Distribution of" ~R^2, 
        subtitle = "Based on 5000 bootstrap samples of a linear model",
        x = "Estimate",
        y = "Density")
```

The estimates for **$\hat{r}^2$** looks to be roughly **normally distributed**. However, it is important to note that the estimate is high, as the values are distributed closer to the upper bound of 1. It could be argued that there is a very slight left-skew in the distribution, but this is likely due to the high estimate of $\hat{r}^2$. This suggests that a vast majority of the variation in `tmax` can be explained by `tmin`.

##### 2.4. Estimate for $\log(\beta_0 * \beta_1)$

Similarly, we plot the distribution of the estimates of $\log(\beta_0 * \beta_1)$, using the code chunk below. Note that we need to do some more data wrangling for this plot.

```{r boostrap_log}
boot_strap %>% 
  ggplot(aes(x = log_b0b1)) + 
  geom_density(alpha = 0.2) +
  stat_function(fun = dnorm,
                args = with(boot_strap, c(mean = mean(log_b0b1), sd = sd(log_b0b1))),
                color = "blue",
                size = 1.5,
                alpha = .7) +
  theme(legend.position = "none") +
  labs(
        title = "Distribution of" ~log(hat(beta)[0] %*% hat(beta)[1]),
        subtitle = "Based on 5000 bootstrap samples of a linear model",
        x = "Estimate",
        y = "Density")
```

Similar to the **$\hat{r}^2$** estimate, the distribution for **$\log(\beta_0 * \beta_1)$** is also roughly **normally distributed**. However, the plot shows that the **peak** of this distribution is not exactly normal as it has a _very small dip_ at the top of the distribution.

#### 2.5. 95% Confidence Intervals

Using the 5000 bootstrap estimates, we can identify the 2.5% and 97.5% quantiles to provide a **95% confidence interval** for **$\hat{r}^2$** and **$\log(\beta_0 * \beta_1)$**.

- The _95% confidence interval_ of **`R-squared`** is (`r round(quantile(pull(boot_strap, r.squared), probs = c(0.025,0.975)), digits = 2)`).

- The _95% confidence interval_ of **`log(beta0 x beta1)`** is (`r round(quantile(pull(boot_strap, log_b0b1), probs = c(0.025,0.975)), digits = 2)`).  

The code chunk below produces a useful table with point estimates and confidence intervals or $\hat{r}^2$ and $\log(\beta_0 * \beta_1)$.

```{r estimate_ci_table}
# Get estimate and confidence interval for each term
tibble(term = c("R-Squared", "Log Product"),
       Estimate = c(mean(pull(boot_strap, r.squared)),
                    mean(pull(boot_strap, log_b0b1))),
       "Lower CI" = c(quantile(pull(boot_strap, r.squared), 0.025),
                    quantile(pull(boot_strap, log_b0b1), 0.025)),
       "Upper CI" = c(quantile(pull(boot_strap, r.squared), 0.975),
                    quantile(pull(boot_strap, log_b0b1), 0.975))) %>%
  knitr::kable()
```

